Getting rid of the \texttt{console.log}:\\Debugging of RxJS-based Applications

### 2020-11-08 slides.tex
Study

\documentclass[aspectratio=169]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usepackage{hyperref}

\usepackage{pgfpages}
\setbeameroption{show notes on second screen}
% \pgfpagesuselayout{2 on 1}[a4paper,border shrink=10mm]

@@

\title[Debugging of RxJS-based Applications]{Getting rid of the \texttt{console.log}:\\Debugging of RxJS-based Applications}
\author{Manuel Alabor, Markus Stolze}
\institute{OST in Rapperswil, Switzerland}
\date{\today}

@@

@@

\begin{frame}[fragile]{Research Questions}
    \begin{enumerate}
        \vfill\item[RQ1] \textbf{What challenges} do software engineers face when debugging RxJS-based applications?
        \vfill\item[RQ2] How can the \textbf{experience} of software engineers during the debugging process of RxJS-based applications \textbf{be improved}?
        \vfill\item[\color{gray}{RQ3}] \color{gray}{What is the \textbf{impact of proposed solutions} on the debugging experience of software engineers?}
    \end{enumerate}
\end{frame}

\note[itemize] {
    \item First: Collect and Validate information
    \item How do engineers debug RxJS?
    \item What do they like? Dislike?
    \item Validate this data in an observational study
    \item Second: Propose a solution to improve
    \item Third: Implement and validate solution
    \item Not part of this work
    \item Already started with in current iteration
}

@@

\begin{frame}[fragile]{Call for ``War Stories''}
    \begin{figure}[H]
        \centering
        \includegraphics[height=0.6\textheight]{figures/tweet-interview.png}
        \caption{\tiny{Source: \url{https://twitter.com/swissmanu/status/1242429409208029185}}}
    \end{figure}
\end{frame}

\note[itemize] {
    \item First: Collect and Validate information
    \item How do engineers debug RxJS?
    \item What do they like? Dislike?
    \item Validate this data in an observational study
    \item Second: Propose a solution to improve
    \item Third: Implement and validate solution
    \item Not part of this work
    \item Already started with in current iteration
}

\begin{frame}[fragile]{Getting Started with RQ1}
	\begin{vfilleditems}
		\item Talk to practitioners
		\item 1:1 Interviews
		\item Written ``War Story'' Reports
	\end{vfilleditems}
\end{frame}

\begin{frame}{Interviews}
	\begin{vfilleditems}
		\item Informal interviews with five participants
	\end{vfilleditems}
\end{frame}

\begin{frame}{War Stories}
	\begin{vfilleditems}
		\item Request via Twitter and direct emails
		\item Responses from five engineers, including one RxJS core team member
	\end{vfilleditems}
\end{frame}

\begin{frame}{Results}

\section{Observational Study}
\begin{frame}{Hypothesis}
    If software engineers must solve an RxJS-based problem, then they \textbf{will instrument the code manually} in order to understand its behavior.
\end{frame}

\begin{frame}{Study Design}
    \begin{vfilleditems}
        \item Observational Study
        \item Two Problems
        \item 25 Minutes per Problem
        \item 
    \end{vfilleditems}
\end{frame}

\begin{frame}{Problem Example}

\begin{frame}{Results}

>>>>

\documentclass[aspectratio=169]{beamer}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{tikz}
\usepackage{hyperref}
\usepackage{pgfplots}
\usepackage{pgfpages}

\pgfplotsset{compat=1.17}
\setbeameroption{show notes on second screen}
% \pgfpagesuselayout{2 on 1}[a4paper,border shrink=10mm]

@@

\title[Debugging of RxJS-based Applications]{Getting rid of the \texttt{console.log}:\\Debugging of RxJS-based Applications}
\author{Manuel Alabor and Markus Stolze}
\institute{OST in Rapperswil, Switzerland}
\date{\today}

@@

\subsection{Why?}

@@

\subsection{Research Questions}
\begin{frame}[fragile]{Research Questions}
    \begin{enumerate}
        \vfill\item[RQ1] \textbf{What challenges} do software engineers face when debugging RxJS-based applications?
        \vfill\item[RQ2] How can the \textbf{experience} of software engineers during the debugging process of RxJS-based applications \textbf{be improved}?
        \vfill\item[\color{gray}{RQ3}] \color{gray}{What is the \textbf{impact of proposed solutions} on the debugging experience of software engineers?}
    \end{enumerate}
\end{frame}

\note[itemize] {
    \item First: Collect and Validate information
    \item Interviews and War Story Reports
    \item Validate this data in an observational study
    \item Second: Propose a solution to improve
    \item Third: Implement and validate solution
    \item Not part of this work
    \item Already started with in current iteration
}

@@

\subsection{Interviews}

\begin{frame}[fragile]{One-to-One Interviews}
	\begin{vfilleditems}
		\item Interviewed five \textbf{Engineers}
		\item How do they \textbf{use} Reactive Programming?
		\item What do they \textbf{like},
		\item ... and what \textbf{dislike} about it?
	\end{vfilleditems}
\end{frame}

\note[itemize] {
    \item We were interested in the overall, ...
    \item the good ...
    \item and bad experiences
    \item Not yet debugging specific
    \item Recruited through personal network
    \item Talked to five engineers
    \item 4 RxJS, 1 Scala Akka Streams
    \item Most knew RxJS through Angular
}

\begin{frame}{Quotes}
    \begin{vfilleditems}
		\item \textit{``[Reactive Programming is] a good way for \textbf{composing} multiple data sources.''}
		\item \textit{``In \textbf{99\%} of all cases, I add \textbf{console.log} statements [\dots] trying to understand what's happening.''}
	\end{vfilleditems}
\end{frame}

\note[itemize] {
    \item All interviewees were overall positive to RP
    \item All value RP for complex data \textbf{dependencies}
    \item and \textbf{transformations}
    \item Concrete Quote about console log
    \item Off the record: ``Debugging RP?''
    \item ... well, then you are f***
}

\subsection{War Stories}
\begin{frame}[fragile]{Call for ``War Stories''}
    \begin{figure}[H]
        \centering
        \includegraphics[height=0.6\textheight]{figures/tweet-interview.png}
        \caption{\tiny{Source: \url{https://twitter.com/swissmanu/status/1242429409208029185}}}
    \end{figure}
\end{frame}

\note[itemize] {
    \item After first "overall" data collection
    \item Interested in specific data about RxJS Debugging
    \item Twitter and direct-email
    \item 5 replies
    \item RxJS core team member
    \item Google developer experts
    \item Fin Tech Engineers
}

\begin{frame}[fragile]{Report Extracts}
	\begin{vfilleditems}
		\item Traditional debugging utilities are \textbf{seldom sufficient}
		\item Tendency to \textbf{manual} code \textbf{modifications}:
		\begin{vfilleditems}
		    \item Add \textbf{trace logs} with \texttt{console.log}
		    \item \textbf{Extract} code parts to \textbf{sand boxed} environments
		\end{vfilleditems}
	\end{vfilleditems}
\end{frame}

\note[itemize] {
    \item Foo
}

\subsection{Insights}

\begin{frame}[fragile]{Key Take Aways}
    \begin{columns}[T]
        \begin{column}{.6\linewidth}
            \vspace{3em}
            \begin{itemize}
                \item The ``\textbf{to hammer a screw}''-problem\bigskip
                \item Tendency to \textbf{manual} code \textbf{modifications}\bigskip
                \item \textbf{Pessimistic} opinions on debugging of RxJS-based applications
	        \end{itemize}
        \end{column}
        \begin{column}{.4\linewidth}
            \begin{figure}
                \centering
                \includegraphics[width=.7\linewidth]{figures/hammer-screw.jpeg}
                \caption{\tiny{Source: Manuel Alabor}}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\note[itemize] {
    \item ... todo
    \item Based on this
    \item Observational study to validate data
    \item Using the following hypothesis:
}

\section{Observational Study}
\subsection{Study Design}

\begin{frame}[fragile]{Hypothesis}
    \begin{quote}
        If software engineers must solve an RxJS-based problem, then they \textbf{will instrument the code manually} in order to understand its behavior.
    \end{quote}
\end{frame}

\note[itemize] {
    \item Foo
}

\begin{frame}{Study Design}
    \begin{vfilleditems}
        \item \textbf{Observational Study}
        \item \textbf{Briefing} upfront for every subject
        \item \textbf{Uncontrolled} Environment
        \item \textbf{Remote} moderation and execution
        \item ``After Action'' \textbf{Survey}
    \end{vfilleditems}
\end{frame}

\note[itemize] {
    \item Remotely conducted observational study
    \item Execution on subjects computer
    \item Individual dev environment
    \item Mimic "Daily dev routine"
    \item Interested in "How would you tackle RxJS debugging"
    \item Two problems, 25 minutes each
    \item ~1h total
    \item Unattended survey afterwards -> RxJS experience data
}

\begin{frame}{Problem 1: Example}
    \begin{columns}[b]
        \begin{column}{.5\linewidth}
            \begin{figure}
                \centering
                \includegraphics[width=.9\linewidth]{figures/problem1-code.png}
                \caption{RxJS Observable}
            \end{figure}
        \end{column}
        \begin{column}{.5\linewidth}
            \begin{figure}
                \centering
                \includegraphics[width=.9\linewidth]{figures/problem1-ui.png}
                \caption{User Interface}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\note[itemize] {
    \item A problem consisted of:
    \item Provided: \textbf{Code}, \textbf{Unit Tests}, \textbf{UI}
    \item Task: Identify and Resolve Bugs
    \item Task: "Speak out loud"
    \item Of course: we did not care IF they actually resolve a problem
    \item Collect as much data how subjects debug
}

\subsection{Study Execution}
\begin{frame}{Subject Population}
    \begin{vfilleditems}
        \item \textbf{Four} professional software engineers
        \item Two currently \textbf{work} with RxJS
        \item Three have \textbf{two or more years} of \textbf{experience} with RxJS
        \item One of them uses RxJS to develop \textbf{backend applications}
    \end{vfilleditems}
\end{frame}

\note[itemize] {
    \item Four engineers recruited from personal network
    \item All use RxJS for frontend development
    \item Only one for backend
    \item Experienced population; most >2y
    \item All: Trace logs
    \item Two: Specialized loggers
    \item Three: Debugger
}

\subsection{Study Results}

\begin{frame}[fragile]{Observed Debugging Techniques}
    \begin{figure}
        \centering
        \begin{tikzpicture}
            \begin{axis}[
                    height=0.6\textheight,
                    width=0.8\linewidth,
                    xbar,
                    xtick={0,1,2,3,4},
                    xlabel={\# of Subjects},
                    symbolic y coords={Add. Tools,Debugger,Trace Logs},
                    ytick=data,
                    enlarge y limits=0.5,
                    enlargelimits=0.15,
                    nodes near coords, nodes near coords align={horizontal}
                ]
                \addplot[
                        fill=pureminimalistic@text@red,
                        draw=pureminimalistic@text@red
                    ]
                    coordinates {
                        (0,Add. Tools) (4,Trace Logs) (2,Debugger) 
                    };
            \end{axis}
        \end{tikzpicture}
    \end{figure}
\end{frame}

\note[itemize] {
    \item All modified/augmented source code manually
    \item -> Trace logs
    \item Two tried traditional debugging
    \item No one extracted code
    \item No one used additional libraries
    \item On told afterwards: Would have been their next step
    \item Based on this evidence: Hypothesis is correct
}

\begin{frame}[fragile]{Interpretation}
    \begin{vfilleditems}
        \item \textbf{Four} professional software engineers
        \item Two currently \textbf{work} with RxJS
        \item Three have \textbf{two or more years} of \textbf{experience} with RxJS
        \item One of them uses RxJS to develop \textbf{backend applications}
    \end{vfilleditems}
\end{frame}

\note[itemize] {
    \item All used manual trace logs
    \item Two tried traditional debugging
    \item No one extracted code
    \item No one used additional libraries
    \item On told afterwards: Would have been their next step
}
